\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-05-20 doc/USER_MANUAL/14_post_processing.tex
updates manual using verbatim for better display in markdown

This script cuts a given portion of the original data, filters it,
transfers the data into a displacement record, and picks the first
P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
without any argument. An example of the usage of the script:
\begin{lyxcode}
{\footnotesize process\_data.pl~-m~CMTSOLUTION~-s~1.0~-l~0/4000~-i~-f~-t~40/500~-p~-x~bp~DATA/1999.330{*}.BH?.SAC}~
\end{lyxcode}
\noindent which has resampled the SAC files to a sampling rate of
1 seconds, cut them between 0 and 4000 seconds, transfered them into
displacement records and filtered them between 40 and 500 seconds,
picked the first P and S arrivals, and added suffix `\texttt{bp}'
to the file names.

@@

This script converts the synthetic output from the SEM code from ASCII
to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
An example of the usage of the script:
\begin{lyxcode}
{\footnotesize process\_syn.pl~-m~CMTSOLUTION~-h~-a~STATIONS~-s~1.0~-l~0/4000~-f~-t~40/500~-p~-x~bp~SEM/{*}.BX?.semd}~
\end{lyxcode}
which will convolve the synthetics with a triangular source-time function
from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
format, add event and station information into the SAC headers, resample
the SAC files with a sampling rate of 1 seconds, cut them between
0 and 4000 seconds, filter them between 40 and 500 seconds with the
same filter used for the observed data, pick the first P and S arrivals,
and add the suffix `\texttt{bp}' to the file names.

@@

The original data and synthetics have three components: vertical (BHZ
resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
for most seismology applications, transverse and radial components
are also desirable. Therefore, we need to rotate the horizontal components
of both the data and the synthetics to the transverse and radial direction,
and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
\begin{lyxcode}
rotate.pl~-l~0~-L~180~-d~DATA/{*}.BHE.SAC.bp~

rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
\end{lyxcode}
where the first command performs rotation on the SAC data obtained
through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
while the second command rotates the processed SEM synthetics.

@@

\begin{lyxcode}
collect\_seismo.pl~machines~DATA/Par\_file
\end{lyxcode}
\end{description}
where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
the parameter file used to extract the \texttt{LOCAL\_PATH} directory
used for the simulation.

@@

After all the simulations are done, the seismograms are collected,
and the useful database files are copied to the frontend, you may
need to clean the local scratch disk for the next simulation. This
is especially important in the case of kernel simulation, where very
large files are generated for the absorbing boundaries to help with
the reconstruction of the regular forward wavefield. A sample script
is provided in \texttt{utils/}:
\begin{lyxcode}
cleanbase.pl~machines
\end{lyxcode}
where \texttt{machines} is a file containing the node names.

@@

With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
code to convert these binary movie data into GMT xyz files for futher
processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
is provided to do this conversion, and then plot the movie snapshots
in GMT, for example:}{\small \par}
\begin{lyxcode}
movie2gif.gmt.pl~-m~CMTSOLUTION~-g~-f~1/40~-n~-2~-p~
\end{lyxcode}
which for the first through the 40th movie frame, converts the \texttt{moviedata}
files into GMT xyz files, interpolates them using the 'nearneighbor'
command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
and `\texttt{-p}' are both optional.

@@

With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
of a shakemap simulation ~\\
 (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
code to convert the binary shakemap data into GMT xyz files. A sample
script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
and then plot the shakemaps in GMT, for example:
\begin{lyxcode}
plot\_shakemap.gmt.pl~data\_dir~type(1,2,3)~CMTSOLUTION~
\end{lyxcode}
where \texttt{type=1} for a displacement shakemap, \texttt{2} for
velocity, and \texttt{3} for acceleration.

@@

A sample program \texttt{remap\_database} is provided to map the local
database from a set of machines to another set of machines. This is
especially useful when you want to run mesher and solver, or different
types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
\begin{lyxcode}
run\_lsf.bash~-{}-gm-no-shmem~-{}-gm-copy-env~remap\_database~old\_machines~150
\end{lyxcode}
where \texttt{old\_machines} is the LSF machine file used in the previous
simulation, and \texttt{150} is the number of processors in total.

>>>>

This script cuts a given portion of the original data, filters it,
transfers the data into a displacement record, and picks the first
P and S arrivals. For more functionality, type `\texttt{process\_data.pl}'
without any argument. An example of the usage of the script:

{\footnotesize
\begin{verbatim}
process_data.pl -m CMTSOLUTION -s 1.0 -l 0/4000 -i -f -t 40/500 -p -x bp DATA/1999.330*.BH?.SAC
\end{verbatim}
}

\noindent 
which has resampled the SAC files to a sampling rate of
1 seconds, cut them between 0 and 4000 seconds, transfered them into
displacement records and filtered them between 40 and 500 seconds,
picked the first P and S arrivals, and added suffix `\texttt{bp}'
to the file names.

@@

This script converts the synthetic output from the SEM code from ASCII
to SAC format, and performs similar operations as `\texttt{process\_data.pl}'.
An example of the usage of the script:

{\footnotesize
\begin{verbatim}
process_syn.pl -m CMTSOLUTION -h -a STATIONS -s 1.0 -l 0/4000 -f -t 40/500 -p -x bp SEM/*.BX?.semd
\end{verbatim}
}
\noindent
which will convolve the synthetics with a triangular source-time function
from the \texttt{CMTSOLUTION} file, convert the synthetics into SAC
format, add event and station information into the SAC headers, resample
the SAC files with a sampling rate of 1 seconds, cut them between
0 and 4000 seconds, filter them between 40 and 500 seconds with the
same filter used for the observed data, pick the first P and S arrivals,
and add the suffix `\texttt{bp}' to the file names.

@@

The original data and synthetics have three components: vertical (BHZ
resp. BXZ), north (BHN resp. BXN) and east (BHE resp. BXE). However,
for most seismology applications, transverse and radial components
are also desirable. Therefore, we need to rotate the horizontal components
of both the data and the synthetics to the transverse and radial direction,
and \texttt{\small rotate.pl}{\small{} can be used to accomplish this:}{\small \par}
\begin{verbatim}
rotate.pl -l 0 -L 180 -d DATA/*.BHE.SAC.bp
rotate.pl -l 0 -L 180 SEM/*.BXE.semd.sac.bp
\end{verbatim}
where the first command performs rotation on the SAC data obtained
through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
while the second command rotates the processed SEM synthetics.

@@

\begin{verbatim}
collect_seismo.pl machines DATA/Par_file
\end{verbatim}

\end{description}

where \texttt{machines} is a file containing the node names and \texttt{DATA/Par\_file}
the parameter file used to extract the \texttt{LOCAL\_PATH} directory
used for the simulation.

@@

After all the simulations are done, the seismograms are collected,
and the useful database files are copied to the frontend, you may
need to clean the local scratch disk for the next simulation. This
is especially important in the case of kernel simulation, where very
large files are generated for the absorbing boundaries to help with
the reconstruction of the regular forward wavefield. A sample script
is provided in \texttt{utils/}:
\begin{verbatim}
cleanbase.pl machines
\end{verbatim}
where \texttt{machines} is a file containing the node names.

@@

With the movie data saved in \texttt{OUTPUT\_FILES/} at the end of
a movie simulation (\texttt{\small MOVIE\_SURFACE=.true.}{\small ),
you can run the }\texttt{\small `create\_movie\_shakemap\_AVS\_DX\_GMT}{\small '
code to convert these binary movie data into GMT xyz files for futher
processing. A sample script }\texttt{\small movie2gif.gmt.pl}{\small{}
is provided to do this conversion, and then plot the movie snapshots
in GMT, for example:}{\small \par}
\begin{verbatim}
movie2gif.gmt.pl -m CMTSOLUTION -g -f 1/40 -n -2 -p
\end{verbatim}
which for the first through the 40th movie frame, converts the \texttt{moviedata}
files into GMT xyz files, interpolates them using the 'nearneighbor'
command in GMT, and plots them on a 2D topography map. Note that `\texttt{-2}'
and `\texttt{-p}' are both optional.

@@

With the shakemap data saved in \texttt{OUTPUT\_FILES/} at the end
of a shakemap simulation ~\\
 (\texttt{CREATE\_SHAKEMAP=.true.}), you can also run \texttt{`create\_movie\_shakemap\_AVS\_DX\_GMT}'
code to convert the binary shakemap data into GMT xyz files. A sample
script \texttt{plot\_shakemap.gmt.pl} is provided to do this conversion,
and then plot the shakemaps in GMT, for example:
\begin{verbatim}
plot_shakemap.gmt.pl data _dir type(1,2,3) CMTSOLUTION
\end{verbatim}
where \texttt{type=1} for a displacement shakemap, \texttt{2} for
velocity, and \texttt{3} for acceleration.

@@

A sample program \texttt{remap\_database} is provided to map the local
database from a set of machines to another set of machines. This is
especially useful when you want to run mesher and solver, or different
types of solvers separately through a scheduler (refer to Chapter~\ref{cha:Scheduler}).
\begin{verbatim}
run_lsf.bash --gm-no-shmem --gm-copy-env remap_database old_machines 150
\end{verbatim}
where \texttt{old\_machines} is the LSF machine file used in the previous
simulation, and \texttt{150} is the number of processors in total.
